### 2015-03-13 ThesisJo/derivations.tex
la2pdf commit - 2015:03:13	14:59:07	(Fri13.Mar)
la2pdf commit - 2015:03:13	15:29:32	(Fri13.Mar)

Now let us choose a trigonometric window function:
%
\begin{align}
\w = e^{j\phi_\theta[m]}\big(\alpha - (1-\alpha)\cos\left(\frac{2\pi m}{M-1}\right)\big)
\end{align}
%
where $\phi_\theta[m] = \frac{2\pi}{\lambda}dm\sin\theta$. Inserting this yields:
%
\begin{align}
\w\H\x &= \sumb{m=0}{M-1} e^{j\phi_\theta[m]}\Big(\alpha - (1-\alpha)\cos\left(\frac{2\pi m}{M-1}\right)\Big) x_m^* \nn
&= \alpha\sumb{m=0}{M-1} e^{j\phi_\theta[m]} x_m^* + (1-\alpha)\sumb{m=0}{M-1} e^{j\varphi[m]}\sin\left(\frac{2\pi m}{M-1}\right) x_m^* \\
&= \alpha \a\T\x + (1-\alpha)\b\T\x
\end{align}
%
where:
%
\begin{align}
\a = \begin{bmatrix}
     1 \\
     e^{j\frac{2\pi}{\lambda}d\sin\theta} \\
     e^{j\frac{2\pi}{\lambda}2d\sin\theta} \\
     \vdots \\
     e^{j\frac{2\pi}{\lambda}d(M-1)\sin\theta}
     \end{bmatrix} \qquad \text{and} \qquad
\b = \begin{bmatrix}
     0 \\
     e^{j\frac{2\pi}{\lambda}d\sin\theta}\sin\left(\frac{2\pi m}{M-1}\right) \\
     e^{j\frac{2\pi}{\lambda}2d\sin\theta}\sin\left(\frac{2\pi m}{M-1}\right) \\
     \vdots \\
     0
     \end{bmatrix}
\end{align}
%
To attain unity gain in the look direction the weights must sum to one. 
%
\begin{align}
\w\T\1 &= \big(\alpha\a - (1-\alpha)\b\big)\T\1 = 1
\end{align}
%
This is true for any value of $\alpha$ if $\a\T\1 = 1$ and $\b\T\1 = 1$.

@@

The frequency response of an element is is given by the Fourier transform of the element's window function.
%
\begin{align}
W(\k) = \int\limits_{V} w(\p)\,e^{-j\k\dot\p} dV
\end{align}
%
where $\p = \bmat{x & y & z}\T$ and $dV = dx\,dy\,dz$. For a rectangular 2D element spanning the xy-plane, having sides with length $d_x$ and $d_y$, the frequency response becomes:
%
\begin{align}
W_\text{rect}(k_x, k_y) = \frac{\sin\frac{k_x\,d_x}{2}}{\frac{k_x\,d_x}{2}}
\end{align}
%
If we assume a rectangular

>>>>

Now let us choose a trigonometric window function:
%
\begin{align}
\w = e^{j\phi_\theta[m]}\big(\alpha - (1-\alpha)\cos\left(\frac{2\pi m}{M-1}\right)\big)
\end{align}
%
where $\phi_\theta[m] = \frac{2\pi}{\lambda}dm\sin\theta$. Inserting this yields:
%
\begin{align}
\w\H\x &= \sumb{m=0}{M-1} e^{j\phi_\theta[m]}\Big(\alpha - (1-\alpha)\cos\left(\frac{2\pi m}{M-1}\right)\Big) x_m^* \nn
&= \alpha\sumb{m=0}{M-1} e^{j\phi_\theta[m]} x_m^* - (1-\alpha)\sumb{m=0}{M-1} e^{j\varphi[m]}\cos\left(\frac{2\pi m}{M-1}\right) x_m^* \\
&= \alpha \a\T\x + (1-\alpha)\b\T\x
\end{align}
%
where:
%
\begin{align}
\a = \begin{bmatrix}
     1 \\
     e^{j\frac{2\pi}{\lambda}d\sin\theta} \\
     e^{j\frac{2\pi}{\lambda}2d\sin\theta} \\
     \vdots \\
     e^{j\frac{2\pi}{\lambda}d(M-1)\sin\theta}
     \end{bmatrix} \qquad \text{and} \qquad
\b = \begin{bmatrix}
     0 \\
     e^{j\frac{2\pi}{\lambda}d\sin\theta}\sin\left(\frac{2\pi m}{M-1}\right) \\
     e^{j\frac{2\pi}{\lambda}2d\sin\theta}\sin\left(\frac{2\pi m}{M-1}\right) \\
     \vdots \\
     0
     \end{bmatrix}
\end{align}
%
To attain unity gain in the look direction the weights must sum to one. 
%
\begin{align}
\w\T\1 &= \big(\alpha\a - (1-\alpha)\b\big)\T\1 = 1
\end{align}
%
This is true for any value of $\alpha$ if $\a\T\1 = 1$ and $\b\T\1 = 1$.

@@

The frequency response of an element is is given by the Fourier transform of the element's window function.
%
\begin{align}
W(\k) = \int\limits_{V} w(\p)\,e^{-j\k\dot\p} dV
\end{align}
%
where $\p = \bmat{x & y & z}\T$ and $dV = dx\,dy\,dz$. For a rectangular 2D element spanning the xy-plane, having sides with length $d_x$ and $d_y$, the frequency response becomes:
%
\begin{align}
W_\text{rect}(k_x, k_y) = \frac{\sin\frac{k_x\,d_x}{2}}{\frac{k_x\,d_x}{2}}\,\frac{\sin\frac{k_y\,d_y}{2}}{\frac{k_y\,d_y}{2}} = \sinc\frac{k_x\,d_x}{2}\,\sinc\frac{k_y\,d_y}{2}
\end{align}

\subsubsection{Opening angle and element size}

What is the element size of the HISAS1030 sonar? This can be inferred from the element frequency response:
%
\begin{align}
W_\text{rect}(k_x) = \frac{\sin\frac{k_x\,d_x}{2}}{\frac{k_x\,d_x}{2}} = \frac{1}{\sqrt{2}}
\end{align}
%
This can be solved for $\frac{k_x\,d_x}{2}$ either by series expansion of $\sin(\cdot)$, or numerically to find that $\frac{k_x\,d_x}{2} \approx 1.39$. For HISAS1030, with an opening angle of 25$^\circ$ at 100kHz in both directions, this means the element size must be quadratic with sides:
%
\begin{align}
d = \frac{2*1.39}{k_x} = \frac{2*1.39}{2\pi \frac{f_c}{c}\sin\theta} = \frac{1.39*1500\text{ms}^{-1}}{\pi\, 100\text{kHz}\,\sin{12.5^\circ}} = 3.07\text{cm}
\end{align}
%
This leaves a little space between each sensor since the element spacing is around 3.75cm.
